---
title: "Descriptive"
author: "Anh Nguyen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
rm(list = ls())
source("utility/setup.r")
head(included.data, 10)
```

# Data Description (Limit to no more than 1 page, excluding tables and figures)

• Describe the data collection and study design.

• Choose and describe the primary and secondary outcomes (at least one continuous and one binary/categorical) and other relevant variables.

• Consider the relationship among the selected variables.

• Evaluate the potential distribution of your outcome variables, using graphical methods and summary statistics. If there is any missing data, they should be described in this section.

• Comment on normal and binomial/multinomial distribution assumptions.


## Relationship - hypotheses: 
- baseline weight: higher at female, african and hispanic american, 
- weight loss:
  - positive correlation between months and weight loss
  - more weight loss are seen at: intervention group 
  - no difference between sites

## References:
https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html


## Describe demographics: Summary statistics
Count the number of records for each categories and show percentages:
e.g.: # of ppl of each gender, education, marital status, tobacco use, alcohol use (binary, use a same table with 3 columns),
  # of ppl of each site (5), race (4)


```{r, demographics}
table1(~ Site + Gender + age + height + Race + comorbid + Education + Marital + Tobacco + Alcohol | Condition, data = included.data, topclass=table1.styles, overall=F, extra.col=list(`P-value`=pvalue))
```
No significant differences in demographics between 2 groups.

```{r}
histogram(~age | Condition, data = included.data, fit = "normal")
histogram(~height | Condition, data = included.data, fit = "normal")
bargraph(~comorbid | Condition, data = included.data)

```
```{r, multinomial distribution}
bargraph(~site | Condition, data = included.data)
bargraph(~Race | Condition, data = included.data)
```

## Describe intervention outcomes: summary statisics and graphical methods
baseline wgt, final wgt, months
histogram/bar chart/scatter , mean, sd


No significant differences in outcome variables between 2 groups
```{r, summary}
table1(~ wgt_base + wgt_final + month | Condition, data = included.data, topclass=table1.styles, overall=F, extra.col=list(`P-value`=pvalue))
```

```{r, histogram}
histogram(~wgt_base | Condition, data = included.data, fit = "normal")
histogram(~wgt_final | Condition, data = included.data, fit = "normal")
histogram(~month | Condition, data = included.data)
```
```{r, boxplots}

ggplot(data = included.data, mapping = aes(x = Condition, y = wgt_base)) +
  plt_theme +
  geom_violin(width = .5) +
  geom_boxplot(
    width = 0.25,
    outlier.size = 1.5) +
  stat_summary(
  fun = "mean",
  geom = "point",
  size = 4,
  shape = 18
  # https://ggplot2.tidyverse.org/articles/ggplot2-specs.html#sec:shape-spec
  ) 
   


```


Maybe trying some QQ plot? => another method to visualize distribution (check normal distribution <=> a perfect straight diagonal line

















)


### Test normality or determine the underlying distribution of variables

Varibles seemingly not normal: month, cormobid

Test normality of: age, height, wgt_base, wgt_final

Use shapiro wilk test

