---
title: "Estimate_CI"
author: "Anh Nguyen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, include=FALSE}
rm(list = ls())
source("utility/setup.r")
head(included.data, 10)
```


# Point Estimates and Confidence Intervals (Limit to no more than 1 page)
• Obtain point estimates for population parameters of interest.
• Obtain confidence intervals for those parameters, using both parametric (e.g., asymptotic) and nonparametric (e.g., bootstrap) methods. Indicate which interval you used. Interpret your results. 

## Some notes
I can construct a table to show the point estimates and CIs of multiple variables.

Which variables? 
- continuous variables (normal distribution): age, height, baseline weight, final weight

- non-normal variables: months, number of comorbid conditions

- binomial distribution: gender, education, tobacco, alcohol, marital status, intervention condition

- multinomial distri: clinical site, race 

Variables that not to be estimated: gender, intervention condition, clinical site, race 

Do I want to separate samples by intervention conditions in this chapter????


<!-- ### Normal Approximation of Binomial distribution -->
<!-- For variables: gender, education, tobacco, alcohol, marital status -->
<!-- Check Rosner book, Chap 5.7 and Lec 3 -->
<!-- ```{r} -->

<!-- ``` -->


## 1. Obtain point estimates for population parameters of interest.
Purpose: model the given data by some probability => as a way to check several assumptions like normality?

What are the params of interest? E.g.: means, variances for normal distribution

Sample mean and sample variance are good point estimates for population mean and variance

Other: MLE (MLE can be incorrect if the assumption about underlying distribution is questionable).


```{r}



```

## 2. Obtain confidence intervals for those parameters, using both parametric (e.g., asymptotic) and nonparametric (e.g., bootstrap) methods. Indicate which interval you used. Interpret your results.

### Asymptotic method
For variables that might follow normal distribution: CI for means and variances
- unknown population mean and variance
Approximation method for finding CIs: CLT,  slutsky's theorem
Using MLE to construct approximate CIs: when the params are unknown

```{r}
list_CI <- list("age", "height", "wgt_base", "wgt_final")

for (label_CI in list_CI){
  print(paste("====",label_CI, "===="))
  print(t.test(x=included.data[[label_CI]], conf.level = 0.95)$conf.int)
  # print("Control")
  # print(t.test(x=included.data[included.data$tx == "0", ][[label_CI]], conf.level = 0.95)$conf.int)
  # print("Intervention")
  # print(t.test(x=included.data[included.data$tx == "1", ][[label_CI]], conf.level = 0.95)$conf.int)
}

```



### Bootstrap method
Used when the the sample are non normal (skewed)

which variables dont seem to be normal? month, comorbid

```{r}

list.bootstrap = list("month", "comorbid")
for (i in list.bootstrap){
  m <- sapply(X = 1:10000,
                    FUN = function(X){mean(sample(x = included.data[[i]], 
                                                  size = nrow(included.data), 
                                                  replace = TRUE))})
  hist(x = m, freq=FALSE, main = paste("Sampling distribution of", i))
  print(i)
  print(quantile(x = m, probs = c(0.025, 0.975)))
}
```




