---
title: "Hypothesis_Test"
author: "Anh Nguyen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, include=FALSE}
rm(list = ls())
source("utility/setup.r")
head(included.data, 10)
```

# Hypothesis Testing (Limit to no more than 1 page)

• Translate the scientific questions of your interest into null and alternative hypotheses.

• Choose appropriate approaches for hypothesis testing. Justify your choice of test and discuss potential issues that you might have in making your choice.

• Carry out the test and interpret your results. Check assumptions if applicable.

## What are the questions?

-   How much weight loss do two groups (intervention and control) achieve? Do intervention group lose more weight than the control group do?

-   Which factors significantly influence the weight loss outcomes and which factors do not?

-   Does intervention duration correlate with the weight loss outcomes?

-   which factors predict better outcome?

## Some Notes

which test is used for testing changes?

Post and pre intervention measurements are dependent =\> paired t-test



Here normality of wgt_base and wgt_final is assumed??? 

[Action] Have to check the assumption of normality or validity for normal approximation 




## Weight changes pre and post intervention

### Are weight losses in 2 group significant?

H1: mean weights at pre and post-intervention are different in both group

Using paired t-test

```{r}
t.test(
  x = included.data[included.data$tx==0,]$wgt_final,
  y = included.data[included.data$tx==0,]$wgt_base,
  paired = TRUE,
  mu = 0,
  alternative = "two.sided",
  conf.level = 0.95
)

t.test(
  x = included.data[included.data$tx==1,]$wgt_final,
  y = included.data[included.data$tx==1,]$wgt_base,
  paired = TRUE,
  mu = 0,
  alternative = "two.sided",
  conf.level = 0.95
)
```
The weight difference in CG is not significant (p=0.072, mean = -0.794)
Intervention group show significant weight loss (mean = -2.48, p = 3.8e-5)


### Do intervention group lose more weight than the control group do?

```{r}
head(included.data$wgt_change, 20)
histogram(~wgt_change| Condition, data = included.data, fit = "normal")
table1(~ wgt_change | Condition, data = included.data, topclass=table1.styles, overall=F, extra.col=list(`P-value`=pvalue))
```

P-value (p=0.00196) here rejects the hypothesis that mean weight losses in 2 groups are equal. Not test if which group lost more weight.

Code below will test :
H0: weight loss in intervention group is no less than in the control group 
vs. 
H1: intervention group lose more weight than the control group do

First, Test equal sample variance between wgt_change in 2 groups
```{r}
var.test(
  x = included.data[included.data$tx==0,]$wgt_change,
  y = included.data[included.data$tx==1,]$wgt_change,
  ratio = 1,
  alternative = "two.sided",
  conf.level = 0.95
)
```
Variance difference in wgt_change between 2 groups is not significant (p=0.113)

Now, assuming equal population variance in wgt_change between 2 groups, we test the difference in mean wgt_change using one.sided test:
```{r}
t.test(
  x = included.data[included.data$tx==1,]$wgt_change,
  y = included.data[included.data$tx==0,]$wgt_change,
  mu = 0,
  var.equal = TRUE,
  alternative = "less",
  conf.level = 0.95
)
```
weight loss in intervention group is significantly more than in CG (p=0.009)


### Which factors significantly influence the weight loss outcomes and which factors do not?

Do this question in Linear model. Hypothesis testing section ends here






